---
title: "MLB Time Series Analysis"
author: "Braden Barglind, Jos Oliveros"
date: '2022-06-13'
output:
  html_document: default
  pdf_document: default
---

### Required Packages

```{r include=FALSE}
library(TSA)
library(sandwich)
library(lmtest)
library(tseries)
library(forecast)
```


#### Time Series Data


```{r}
overall.data <- read.csv('MLB Stats - Overall.csv')
overall.data <- subset(overall.data,select= c(-X))
overall.data <- ts(overall.data, start=1961)

plot(overall.data,type='o', main="Statistics Over Time in MLB")
```

#### Train/Test Split

```{r}
train <- subset(overall.data, subset= time(overall.data) < 2010)
test <- subset(overall.data, subset= time(overall.data) >= 2010)

#split data for testing true model
train <- ts(train, start=1961)
test <- ts(test, start=2010)

plot(train,type='o')
```

 

```{r}
HR.data <- train[,"HR"]
hr.model <- lm(HR.data~time(HR.data))
loghr.model <- lm(log(HR.data)~time(HR.data))

summary(hr.model)
summary(loghr.model)
```



```{r}
hits.data <- train[,"Hits"]
hits.model <- lm(hits.data~time(hits.data))
sqrthits.model <- lm(sqrt(hits.data)~time(hits.data))

summary(hits.model)
summary(sqrthits.model)
```
 

```{r}
runs.data <- train[,"Runs"]
runs.model <- lm(runs.data~time(runs.data))
logruns.model <- lm(log(runs.data)~time(runs.data))
sqrtruns.model <- lm(sqrt(runs.data)~time(runs.data))


summary(runs.model)
summary(logruns.model)
summary(sqrtruns.model)
```

```{r}
SO.data <- train[,"SO"]
so.model <- lm(SO.data~time(SO.data))
logso.model <- lm(log(SO.data)~time(SO.data))

summary(so.model)
summary(logso.model)
```

### Best Fit Line

```{r}
par(mfrow=c(2,2))
plot(HR.data,type='o');abline(hr.model)
plot(hits.data,type='o');abline(hits.model)
plot(runs.data,type='o');abline(runs.model)
plot(SO.data,type='o');abline(so.model)
```

#### HAC TEST

```{r}
 
coeftest(hr.model, vcov=vcovHAC(hr.model))
coeftest(hits.model, vcov=vcovHAC(hits.model))
coeftest(runs.model, vcov=vcovHAC(runs.model))
coeftest(so.model, vcov=vcovHAC(so.model))
```


### Non-transformed model residuals

```{r}
par(mfrow=c(2,2))
plot(rstandard(hr.model),type='o')
plot(rstandard(hits.model),type='o')
plot(rstandard(runs.model),type='o')
plot(rstandard(so.model),type='o')

```

### Transformed Model Residuals

```{r}
hrres <- rstandard(hr.model)
hitsres <- rstandard(sqrthits.model)
runsres <- rstandard(sqrtruns.model)
sores <- rstandard(logso.model)

par(mfrow= c(2,2))
plot(hrres,type='o')
plot(hitsres,type='o')
plot(runsres,type='o')
plot(sores,type='o')
```

```{r}
par(mfrow=c(2,2))
qqnorm(hrres);qqline(hrres)
qqnorm(hitsres);qqline(hitsres)
qqnorm(runsres);qqline(runsres)
qqnorm(sores);qqline(sores)
```

```{r}
par(mfrow=c(2,2))
hist(hrres)
hist(hitsres)
hist(runsres)
hist(sores)
```


```{r}
shapiro.test(hrres)
shapiro.test(hitsres)
shapiro.test(runsres)
shapiro.test(sores)
```




### Time Dependence



```{r}
par(mfrow=c(2,2))
acf(hrres)
acf(hitsres)
acf(runsres)
acf(sores)
```



```{r}
par(mfrow=c(2,2))
runs(hrres)
runs(hitsres)
runs(runsres)
runs(sores)
```


### Determine Order of Integration d



```{r}
par(mfrow=c(2,2))
plot(hrres, type='o')
plot(hitsres, type='o')
plot(runsres, type='o')
plot(sores, type='o')
```

#### Stationarity and Unit Roots

```{r}
adf.test(hrres)
adf.test(hitsres)
adf.test(runsres)
adf.test(sores)
```

```{r}
pp.test(hrres)
pp.test(hitsres)
pp.test(runsres)
pp.test(sores)
```

```{r}
kpss.test(hrres)
kpss.test(hitsres)
kpss.test(runsres)
kpss.test(sores)
```

### Choosing p and q

```{r}
par(mfrow=c(2,2))
acf(hrres)
acf(hitsres)
acf(runsres)
acf(sores)
```

```{r}
par(mfrow=c(2,2))
pacf(hrres)
pacf(hitsres)
pacf(runsres)
pacf(sores)
```

#### EACF Plots

```{r}
print("HR")
eacf(hrres)
print("Hits")
eacf(hitsres)
print("Runs")
eacf(runsres)
print("SO")
eacf(sores)

```

For HR:
IMA(1,1) or ARI(3,1)

For Hits:
IMA(1,1) or ARI(3,1)

For Runs:

ARI(1,1)

For SO:

ARI(5,1) or IMA(1,1)



### Creating ARIMA Models

```{r}
hr.MA <- Arima(hrres, c(0,1,1))
hr.AR3 <- Arima(hrres, c(3,1,0))
hr.MA
hr.AR3
```

```{r}
hits.MA <- Arima(hitsres, c(0,1,1))
hits.AR <- Arima(hitsres, c(3,1,0))
hits.MA
hits.AR
```

```{r}
runs.AR <- Arima(runsres, c(1,1,0))
runs.AR
```

```{r}
so.AR <- Arima(sores, c(5,1,0))
so.MA <- Arima(sores, c(0,1,1))
so.AR
so.MA
```

keep `hr.MA`, `hits.MA`, `runs.AR`, `so.MA`

```{r}
tsdiag(hr.MA)
tsdiag(hits.MA)
tsdiag(runs.AR)
tsdiag(so.MA)
```

### Overfitting

```{r}
hr.overfit <- Arima(hrres, c(0,1,2))
hits.overfit <- Arima(hitsres, c(0,1,2))
runs.overfit <- Arima(runsres, c(2,1,0))
so.overfit <- Arima(sores, c(0,1,2))
```

```{r}
data.frame(
  Model= c("hr.MA","hits.MA","runs.AR","so.MA","hr.overfit","hits.overfit","runs.overfit","so.overfit"),
  AIC= c(hr.MA$aic, hits.MA$aic, runs.AR$aic, so.MA$aic, hr.overfit$aic, hits.overfit$aic, runs.overfit$aic, so.overfit$aic),
  AICC= c(hr.MA$aicc, hits.MA$aicc, runs.AR$aicc, so.MA$aicc, hr.overfit$aicc, hits.overfit$aicc, runs.overfit$aicc, so.overfit$aicc),
  BIC= c(hr.MA$bic, hits.MA$bic, runs.AR$bic, so.MA$bic, hr.overfit$bic, hits.overfit$bic, runs.overfit$bic, so.overfit$bic)
  )
```

### Forecasting

```{r}
#HR
HR.MA1_xreg=Arima(HR.data,order=c(0,1,1),xreg=1:length(HR.data))
newtm=seq(from=length(HR.data)+1,to=length(HR.data)+10,length=10)
predx=predict(HR.MA1_xreg,n.ahead=10,newxreg=newtm)
pr=predx$pred
uci=pr+2*predx$se
lci=pr-2*predx$se

# To plot the predicted values as prediction intervals, code them as time series
pr=ts(pr,start=2010,freq=1)
uci=ts(uci,start=2010,freq=1)
lci=ts(lci,start=2010,freq=1)

ymin=min(c(as.vector(lci),HR.data))-.1
ymax=max(c(as.vector(uci),HR.data))+.1

par(mfrow=c(1,1))
plot(HR.data,xlim=c(1961,2020),ylim=c(ymin,ymax),main="HR.data")
lines(test[,"HR"], type='o')
lines(pr,col=2)
lines(uci,col=3)
lines(lci,col=3)
```

```{r}
#Hits
hits.MA1_xreg=Arima(hits.data,order=c(0,1,1),xreg=1:length(hits.data))
newtm=seq(from=length(hits.data)+1,to=length(hits.data)+10,length=10)
predx=predict(hits.MA1_xreg,n.ahead=10,newxreg=newtm)
pr=predx$pred
uci=pr+2*predx$se
lci=pr-2*predx$se

# To plot the predicted values as prediction intervals, code them as time series
pr=ts(pr,start=2010,freq=1)
uci=ts(uci,start=2010,freq=1)
lci=ts(lci,start=2010,freq=1)

ymin=min(c(as.vector(lci),hits.data))-.1
ymax=max(c(as.vector(uci),hits.data))+.1

par(mfrow=c(1,1))
plot(hits.data,xlim=c(1961,2020),ylim=c(ymin,ymax),main="Hits.data")
lines(test[,"Hits"], type='o')
lines(pr,col=2)
lines(uci,col=3)
lines(lci,col=3)
```

```{r}
#Runs
runs.AR1_xreg=Arima(runs.data,order=c(1,1,0),xreg=1:length(runs.data))
newtm=seq(from=length(runs.data)+1,to=length(runs.data)+10,length=10)
predx=predict(runs.AR1_xreg,n.ahead=10,newxreg=newtm)
pr=predx$pred
uci=pr+2*predx$se
lci=pr-2*predx$se

# To plot the predicted values as prediction intervals, code them as time series
pr=ts(pr,start=2010,freq=1)
uci=ts(uci,start=2010,freq=1)
lci=ts(lci,start=2010,freq=1)

ymin=min(c(as.vector(lci),runs.data))-.1
ymax=max(c(as.vector(uci),runs.data))+.1

par(mfrow=c(1,1))
plot(runs.data,xlim=c(1961,2020),ylim=c(ymin,ymax),main="runs.data")
lines(test[,"Runs"], type='o')
lines(pr,col=2)
lines(uci,col=3)
lines(lci,col=3)
```

```{r}
#SO
SO.MA1_xreg=Arima(SO.data,order=c(0,1,1),xreg=1:length(SO.data))
newtm=seq(from=length(SO.data)+1,to=length(SO.data)+10,length=10)
predx=predict(SO.MA1_xreg,n.ahead=10,newxreg=newtm)
pr=predx$pred
uci=pr+2*predx$se
lci=pr-2*predx$se

# To plot the predicted values as prediction intervals, code them as time series
pr=ts(pr,start=2010,freq=1)
uci=ts(uci,start=2010,freq=1)
lci=ts(lci,start=2010,freq=1)

ymin=min(c(as.vector(lci),SO.data))-.1
ymax=max(c(as.vector(uci),SO.data))+.1

par(mfrow=c(1,1))
plot(SO.data,xlim=c(1961,2020),ylim=c(ymin,ymax),main="SO.data")
lines(test[,"SO"], type='o')
lines(pr,col=2)
lines(uci,col=3)
lines(lci,col=3)
```
